Titeln på avhandlingens språk

### 2022-11-07 conclusion.tex
tweaking
too tired to continue

The study and benchmarking of \ac{WCA} applications is a challenging discipline due to these application's intrinsic human-in-the-loop nature.
Humans are notoriously unreliable, and greatly complicate the scalability and repeatability of experiments.
Furthermore, recruiting large enough cohorts of humans for large-scale experimentation is both greatly time-consuming and prohibitively expensive for many research groups.
In this paper, we have presented the \edgedroid{} model of human timing behavior for \ac{WCA}

>>>>

The study and benchmarking of \ac{WCA} applications is a challenging discipline due to these application's intrinsic human-in-the-loop nature.
Humans are notoriously unreliable, and greatly complicate the scalability and repeatability of experiments.
Furthermore, recruiting large enough cohorts of humans for large-scale experimentation is both greatly time-consuming and prohibitively expensive for many research groups.

In the first half of this paper, we have introduced the \edgedroid{} model of human timing behavior for \ac{WCA}, the first data-driven model for human timings in \ac{WCA} applications.
This model represents a stochastic approach to execution time modeling which builds upon the data collected for \textcite{olguinmunoz:impact2021}.
Together with this model, we have also introduced a novel procedure for the generation of synthetic traces of frames in step-based \ac{WCA}, allowing for a full end-to-end emulation of a human when combined with the timing model.

In the second half, we have explored the potential for optimization in \ac{WCA} systems using the previously discussed timing models.
We have proposed a novel stochastic optimization framework for resource consumption-system responsiveness trade-offs in \ac{WCA}.
We have shown that this framework is applicable to a myriad of parameters in these applications, and showcased experimental results employing this framework for the minimization of number of samples processed per step and total energy consumption per step.
Our results show up to a \SI{50}{\percent} increase in performance with respect to state of the art when optimizing for number of samples, and up to a \SI{30}{\percent} improvement when optimizing for energy consumption, proving thus the value of such frameworks for the design of \ac{WCA} applications.

\todo[inline]{Finish}

### 2022-11-07 implications_optimization.tex
too tired to continue

In the following, we will show the effects of our optimization approach combined with our timing models compared to \textcite{Wang2019Towards}'s state-of-the-art approach.
For this we implement these sampling schemes in Python and run a number of simulations with them.
The first scheme uses our approach and \cref{eq:epsilon_terminal,eq:optimization:sampling} to determine the optimum sampling instants and the values for \( \alpha \) and \( \beta \) at each step.
It includes an embedded timing model (without any distribution fitting) to provide updated estimates of the mean execution time \( \mu \) and \( \sigma \) at every step as well, allowing it to adapt to the state of the system.
We will refer to this scheme as the \emph{sample-count-optimized aperiodic} sampling scheme.

@@

We proceed to set up an experiment where these sampling approaches are deployed on identical, simulated, tasks with varying constant \acp{RTT}.
Experimental parameters are summarized in \cref{tab:params}.
We set the \( w_0 \) and \( \beta \) factors of our sample-count-optimized scheme to \SI{1.0}{\second} and \num{1.0}, respectively, for mathematical simplicity, and derive \( \alpha = 1.9 \sigma^{-1} \).
Next, we derive \( R_\text{min} \) for \textcite{Wang2019Towards}'s scheme from \( w_0 \).
As discussed in \cref{sec:background}, we assume that wait times are uniformly distributed between \SI{0}{\second} and sampling interval of each step.
A maximum expected wait time \( w_0 = 1.0\,\si{\second} \) thus translates into a maximum expected sampling interval of \SI{2.0}{\second}, yielding a minimum sampling rate \( R_\text{min} = {2.0\,\si{\second}}^{-1} = 0.5\,\si{\hertz} \).

@@

\begin{figure*}
    \centering
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/new_model/sampling_optimization.png}
        \caption{%
            Round-trip time versus mean number of captured samples per step, averaged over all \num{100} repetitions of the experiment.
            Note the logarithmic scale on the vertical axis.
            Error bars indicate \SI{95}{\percent} \acp{CI}.
        }
    \end{subfigure}\\
    \medskip
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/new_model/sampling_optimization_diff.png}
        \caption{%
            Percentage difference in mean number of captured samples per step with respect to the greedy sampling scheme.
            Error bars indicate \SI{95}{\percent} \acp{CI}, calculated using a two-sided t-test.
        }
    \end{subfigure}
    \caption{%
        Summary of results for experiment comparing the sample-count-optimized aperiodic sampling scheme to the reference schemes and \textcite{Wang2019Towards}'s \ac{CDF}-based approach.
    }\label{fig:optimization:samples}
\end{figure*}

The results of this investigation are presented in \cref{fig:optimization:samples}.
\todo[inline]{Discuss results.}

@@

\begin{figure*}
    \centering
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/new_model/energy_optimization.png}
        \caption{%
            Round-trip time versus mean per step energy consumption, averaged over all \num{100} repetitions of the experiment.
            Error bars indicate \SI{95}{\percent} \acp{CI}.
        }
    \end{subfigure}\\
    \medskip
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/new_model/energy_optimization_diff.png}
        \caption{%
            Percentage difference in mean per step energy consumption with respect to the greedy sampling scheme.
            Error bars indicate \SI{95}{\percent} \acp{CI}, calculated using a two-sided t-test.
        }
    \end{subfigure}
    \caption{%
        Summary of results for experiment comparing the energy-optimized aperiodic sampling scheme to the reference schemes and \textcite{Wang2019Towards}'s \ac{CDF}-based approach.
    }\label{fig:optimization:energy}
\end{figure*}

Again, we include the reference greedy and offline optimum schemes, as well as \textcite{Wang2019Towards}'s approach, and repeat the experiment \num{100} times for each combination of sampling scheme, neuroticism, and \ac{RTT}.
The results of these experiments are presented in \cref{fig:optimization:energy}.

\todo[inline]{Discussss}

>>>>

In the following, we will show the effects of our optimization approach combined with our timing models compared to \textcite{Wang2019Towards}'s state-of-the-art approach.
For this we implement these sampling schemes in Python and run a number of simulations with them.
The first scheme uses our approach, \cref{eq:epsilon_terminal,eq:optimization:sampling}, to determine the optimum sampling instants and the values for \( \alpha \) and \( \beta \) at each step.
It includes an embedded timing model (without any distribution fitting) to provide updated estimates of the mean execution time \( \mu \) and \( \sigma \) at every step as well, allowing it to adapt to the state of the system.
We will refer to this scheme as the \emph{sample-count-optimized aperiodic} sampling scheme.

@@

We proceed to set up an experiment where these sampling approaches are deployed on identical, simulated, tasks with varying constant \acp{RTT}.
Experimental parameters are summarized in \cref{tab:params}.
We set the \( w_0 \) and \( \beta \) factors of our sample-count-optimized scheme to \SI{1.0}{\second} and \num{1.0}, respectively, for mathematical simplicity, and derive \( \alpha = 1.9 \sigma^{-1} \).
Next, we derive \( R_\text{min} \) for \textcite{Wang2019Towards}'s scheme from \( w_0 \).
As discussed in \cref{sec:background}, we assume that wait times are uniformly distributed between \SI{0}{\second} and sampling interval of each step.
A maximum expected wait time \( w_0 = 1.0\,\si{\second} \) thus translates into a maximum expected sampling interval of \SI{2.0}{\second}, yielding a minimum sampling rate \( R_\text{min} = {2.0\,\si{\second}}^{-1} = 0.5\,\si{\hertz} \).
It should also be noted that in both aperiodic schemes, our sampling-count-optimized approach and the \ac{CDF}-based approach, there exists the possibility for sampling instants to be \emph{missed} due to the actual \ac{RTT} of the system being higher than the parameterization of the schemes.
In these cases, both schemes will degrade into greedy sampling as they attempt to sample as fast as possible.

@@

\begin{figure*}
    \centering
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/new_model/sampling_optimization.png}
        \caption{%
            Round-trip time versus mean number of captured samples per step, averaged over all \num{100} repetitions of the experiment.
            Note the logarithmic scale on the vertical axis.
            Error bars indicate \SI{95}{\percent} \acp{CI}.
        }
    \end{subfigure}\\
    \medskip
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/new_model/sampling_optimization_diff.png}
        \caption{%
            Percentage difference in mean number of captured samples per step by the sample-count-optimized approach with respect to the three reference schemes.
            In other words, curves represent the relative performance of the sample-count-optimized scheme when using the corresponding reference scheme as baseline.
            Error bars indicate \SI{95}{\percent} \acp{CI}, calculated using a two-sided t-test.
        }
    \end{subfigure}
    \caption{%
        Summary of results for experiment comparing the sample-count-optimized aperiodic sampling scheme to the reference schemes and \textcite{Wang2019Towards}'s \ac{CDF}-based approach.
    }\label{fig:optimization:samples}
\end{figure*}

The results of this investigation are presented in \cref{fig:optimization:samples}, and clearly show the advantages of using the sample-count-optimized scheme over the current state-of-the-art.
The performance of \textcite{Wang2019Towards}'s approach appears to degrade with lower \acp{RTT}, exponentially oversampling as latency tends to zero and the maximum sampling rate of the system tends to infinity.
On the other hand our sampling scheme exhibits consistently matches or beats the state-of-the-art while maintaining a relatively constant behavior with respect to \acp{RTT}.
This makes the sample-count-optimized approach more predictable, reliable, and consistent, and thus more suitable for real-world deployments.

It should be noted that although both schemes seem to tend towards \num{2.0} samples per step as \acp{RTT} increase, this is simply an artifact of our experimental setup.
As \acp{RTT} increase way above the expected wait time, the probability of the sum of the first sampling interval and the \ac{RTT} being larger than the execution time of the step tends towards \num{1.0}.
This leads to these sampling schemes consistently sampling only twice each step: a first sample which is taken before the execution time of the step, and a second one \ac{RTT} seconds later, after the execution time has been reached.

@@

\begin{figure*}
    \centering
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/new_model/energy_optimization.png}
        \caption{%
            Round-trip time versus mean per step energy consumption, averaged over all \num{100} repetitions of the experiment.
            Error bars indicate \SI{95}{\percent} \acp{CI}.
        }
    \end{subfigure}\\
    \medskip
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/new_model/energy_optimization_diff.png}
        \caption{%
            Percentage difference in mean per step energy consumption by the energy-optimized sampling scheme with respect to the three reference schemes.
            Error bars indicate \SI{95}{\percent} \acp{CI}, calculated using a two-sided t-test.
        }
    \end{subfigure}
    \caption{%
        Summary of results for experiment comparing the energy-optimized aperiodic sampling scheme to the reference schemes and \textcite{Wang2019Towards}'s \ac{CDF}-based approach.
    }\label{fig:optimization:energy}
\end{figure*}

Again, we include the reference greedy and offline optimum schemes, as well as \textcite{Wang2019Towards}'s approach, and repeat the experiment \num{100} times for each combination of sampling scheme, neuroticism, and \ac{RTT}.

The results of these experiments are presented in \cref{fig:optimization:energy}, and they clearly illustrate the advantages of the integration of our timing models with an adaptive, energy-optimized sampling scheme when compared to unoptimized and state-of-the-art sampling schemes.
Our approach is consistently consumes \SI{20}{\percent} less energy than \textcite{Wang2019Towards}'s state-of-the-art, and is up to \SI{40}{\percent} more efficient than the greedy scheme.
Furthermore, once again the behavior of our approach is more consistent and reliable than the competition, exhibiting a flat curve of energy consumption much akin to that of the offline optimum in behavior.
